Peaks Over Threshold for Bursty Time Series

### 2018-04-18 article_springer/article_springer.Rmd
created a bunch of time prediction plots

Figures \ref{fig:flare-diagnostics-1} and \ref{fig:flare-diagnostics-2} 
show the diagnostic plots for a 
minimum threshold chosen at the 200th order statistic. 
There is some residual autocorrelation for the sequence of threshold 
exceedance times that is not accounted for by the MRP model. 
The fit with a Mittag-Leffler distribution ($\beta = 0.8$) is good, and 
there is no apparent dependence between threshold exceedance times and 
event magnitudes.

>>>>

Figures \ref{fig:flare-diagnostics-1} and \ref{fig:flare-diagnostics-2} 
show the diagnostic plots for a 
minimum threshold chosen at the 200th order statistic. 
There is some residual autocorrelation for the sequence of threshold 
exceedance times that is not accounted for by the MRP model. 
The fit with a Mittag-Leffler distribution ($\beta = 0.8$) is good, 
though there are signs that the power-law tail tapers off for very large 
inter-threshold crossing times. 
There is no apparent dependence between threshold exceedance times and 
event magnitudes seen in the copula plot.

# Predicting the time of the next threshold crossing

According to Figure \ref{fig:flares}, for a threshold $\ell$ at the $k$-th order 
statistic, the fitted threshold exceedance time distribution is 
$$
T_\ell \sim {\rm ML}(\beta, k^{1/\beta} \sigma_0), 
$$
where $\beta = 0.85$ and $\sigma_0 = 3.0 \times 10^7 {\rm sec}$. 
Unlike the exponential distribution, the Mittag-Leffler distribution is not 
memoryless, and the probability density of the time $t$ until the next threshold 
crossing will depend on the time $t_0$ elapsed since the last threshold crossing.
This density equals 
$$
p(t|\beta, \sigma_0, \ell, t_0) = \frac{f(t + t_0 | \beta, k^{1/\beta} \sigma_0)}{\mathbf P[T_\ell > t_0]}
$$
where $f(\,\cdot\, | \beta, k^{1/\beta} \sigma_0)$ is the probability density of 
${\rm ML}(\beta, k^{1/\beta} \sigma_0)$.

Conditional on the last threshold being

```{r, fig.height=7}
tail <- 0.7
scale <- 1
t_0 <- c(0, 1, 10)
from <- 0.1
to <- 100 * scale
xx <- exp(seq(from = log(from), to = log(to), length.out = 100))
par(mfrow = c(2,2))
plot(xx, rep(1, length(xx)), ylim = c(0.001,1), type = 'n', log = 'xy', 
     xlab = 't', ylab = 'p')
for (k in 1:length(t_0)) {
  cond <- pml(q = t_0[k], tail = tail, scale = scale, lower.tail = FALSE)
  yy <- dml(x = xx + t_0[k], tail = tail, scale = scale) / cond
  lines(xx,yy, col = k+1, lty = k)
}
legend("topright", c("t_0 = 0", "t_0 = 1", "t_0 = 10"), col = 2:4, lty = 1:3)

plot(xx, rep(1, length(xx)), ylim = c(0.001,1), type = 'n', log = 'xy', 
     xlab = 't', ylab = 'p')
betas <- c(0.8, 0.95, 0.99)
t_0 <- 1
for (k in 1:length(betas)) {
  cond <- pml(q = t_0, tail = betas[k], scale = scale, lower.tail = FALSE)
  yy <- dml(x = xx + t_0, tail = betas[k], scale = scale) / cond
  lines(xx, yy, col = k+1, lty = k)
}
legend("topright", c("beta = 0.8", "beta = 0.95", "beta = 0.99"), col = 2:4, lty = 1:3)
```

```{r}
betas <- c(0.7, 0.8, 0.95, 0.99)
t_0 <- 1
plot(xx, rep(1, length(xx)), ylim = c(0.001,1), type = 'n', log = 'xy', 
     xlab = 't', ylab = 'p')
for (k in 1:length(betas)) {
  yy <- dml(x = xx, tail = betas[k], scale = scale) / pml(
    q = xx,
    tail = betas[k],
    scale = scale,
    lower.tail = FALSE
  )
  lines(xx,yy, col = k+1, lty = k)
}
legend("topright", c("beta = 0.8", "beta = 0.95", "beta = 0.99"), col = 2:5, lty = 1:4)

```
